### 2020-09-03 code/R_markdown/P5_mirzapur_analysis_figures.Rmd

---
title: "Microbiome analysis - Mirzapur follow-up"
author: "Maureen A. Carey"
date: "09/01/2020"
output: html_document
---

# Load packages and data

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(ggplot2); library(phyloseq)

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P5output_plot_2020_09_02.RData")

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000); theme_set(theme_bw())
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

```

# Plot - Fig 4F

```{r}

ggplot(df, aes(x = Site, y = value, color = Site)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitter(width = 0.3, height = 0, seed = SEED_INT)) + 
  facet_wrap(~time) +
  scale_y_sqrt() +
  scale_color_manual(values = c('black','darkgray'), guide = F) +
  geom_hline(yintercept = 0.2, linetype = 'dotted', alpha = .75) +
  ylab('relative abundance')
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4F_mega_box_Mirzapur.png",
       width = 3.2,height = 3.5,units = "in",dpi = 300)

```

## Table 1-like data for methods

```{r}

length(unique(sample_data(ps_mirzapur_rare)$Child.ID))
length((sample_data(ps_mirzapur_rare)$Child.ID))
table(sample_data(ps_mirzapur_rare)$Clinical.type)
min(sample_data(ps_mirzapur_rare)$Date.of.collection)
max(sample_data(ps_mirzapur_rare)$Date.of.collection)

```