### 2020-09-14 code/R_markdown/P5_mirzapur_analysis.Rmd

---
title: "Microbiome analysis - Mirzapur follow-up"
author: "Maureen A. Carey"
date: "09/01/2020"
output: html_document
---

# Load packages and data

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(phyloseq); library(reshape2)
library(rmarkdown); library(dplyr); library(tidyr); library(tibble)

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
rm(list= ls()[! (ls() %in% c('left_study_ids'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p2_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p3_2020_09_11.RData")
ps_mirzapur = readRDS("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/ps_2020_08_12.rds")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p1_2020_09_11.RData") # load save_df1 and save_df2

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000)
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

```

# Check Megasphaera in Mirzapur

```{r}

ps_mirzapur = subset_samples(ps_mirzapur, Site == 2) # MIRZAPUR ONLY

# Quality control - remove mitochondrial and chloroplast reads
og_taxa = data.frame(tax_table(ps_mirzapur))

# first remove non-bacteria
keep = rownames(og_taxa[og_taxa$Kingdom %in% c("Bacteria","Archaea"),])
ps_mirzapur = prune_taxa(keep,ps_mirzapur)
keep = rownames(og_taxa[og_taxa$Family != "Mitochondria",])
ps_mirzapur = prune_taxa(keep,ps_mirzapur)
keep = rownames(og_taxa[og_taxa$Class != "Chloroplast",])
ps_mirzapur = prune_taxa(keep,ps_mirzapur)

# remove if error in any reporting re Crypto infection
keep1 = ifelse(sample_data(ps_mirzapur)$Crypto..Any. == '1' & as.numeric(sample_data(ps_mirzapur)$Clinical.type) < 3, 0, 1)
ps_mirzapur = prune_samples(as.logical(keep1),ps_mirzapur)
keep2 = ifelse(sample_data(ps_mirzapur)$Crypto..Any. == '999' & as.numeric(sample_data(ps_mirzapur)$Clinical.type) > 3, 0, 1)
ps_mirzapur = prune_samples(as.logical(keep2),ps_mirzapur)
keep3 = ifelse(sample_data(ps_mirzapur)$Crypto..Any. == '888' & as.numeric(sample_data(ps_mirzapur)$Clinical.type) > 3, 0, 1)
ps_mirzapur = prune_samples(as.logical(keep3),ps_mirzapur)
rm(keep, keep1, keep2, keep3)

# remove children who were lost to follow up
ps_mirzapur = subset_samples(ps_mirzapur, !(Child.ID %in% left_study_ids))

# Rarify reads in all samples
rare_min = 10000
# remove samples with fewer counts than the minimum
ps_mirzapur_rare = prune_samples(sample_sums(ps_mirzapur) >=rare_min,ps_mirzapur) #8 removed
# remove samples with NA age
ps_mirzapur_rare = prune_samples(!is.na(sample_data(ps_mirzapur_rare)[,"Age.in.days"])[,1],ps_mirzapur_rare) # none removed
ps_mirzapur_rare = rarefy_even_depth(ps_mirzapur_rare,sample.size=rare_min,rngseed=SEED_INT) # OTUs no longer in the dataset will be removed

ntaxa(subset_taxa(ps_mirzapur_rare, Genus == "Megasphaera"))

```

# Extract ASV that important from classifiers

```{r}

otu_for_comp_TOD = as.data.frame(otu_table(ps_mirzapur_rare))[,save_df1$variable]
otu_for_comp_PD = as.data.frame(otu_table(ps_mirzapur_rare))[,save_df1$variable]

# get TOD values
plt_df = melt(otu_for_comp_TOD %>% rownames_to_column('sample')) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() 
tax = tax_table(ps_mirzapur_rare)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
twoASVs_TOD = tax2 %>% filter(var == tax2$var[grep("Megasphaera", tax2$var)])

# get PD values
plt_df = melt(otu_for_comp_PD %>% rownames_to_column('sample')) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() 
tax = tax_table(ps_mirzapur_rare)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
PD_twoASVs = tax2 %>% filter(var == tax2$var[grep("Megasphaera", tax2$var)])

twoASVs = rbind(twoASVs_TOD %>% mutate(time = 'TOD'), PD_twoASVs %>% mutate(time = 'PD'))
twoASVs_wide = tidyr::spread(twoASVs %>% select(value_use,time,sample, Genus),
                             key = Genus, value = value_use) %>% 
  mutate(time = as.factor(time))
levels(twoASVs_wide$time)[levels(twoASVs_wide$time)=="PD"] =  "pre-detection"
levels(twoASVs_wide$time)[levels(twoASVs_wide$time)=="TOD"] =  "time-of-detection"

```

# Prep Plot - Fig 4F

```{r}

count_df_PD_mirzapur = twoASVs_wide %>% filter(time == 'pre-detection')
count_df_TOD_mirzapur = twoASVs_wide %>% filter(time != 'pre-detection')
df = data.frame(
  Site = c('Mirzapur','Mirpur'),
  Megasphaera = c(round(sum(count_df_PD_mirzapur$Megasphaera>0)/length(count_df_PD_mirzapur$Megasphaera),4)*100, round(sum(count_df_PD_mirpur$Megasphaera>0)/length(count_df_PD_mirpur$Megasphaera),4)*100))
df = melt(df)
# ggplot(df) + 
#   geom_bar(aes(x = Site, y = value, fill = Site), stat = 'identity', position = "dodge") + 
#   scale_fill_manual(values = c('black','grey'), guide = F) +
#   ylab('Percent of samples with reads') + xlab(NULL)

#UPDATE HERE: count_df_PD_mirpur column names have switched from genus to ASV
count_df_PD_mirzapur
df_mirz = melt(rbind(count_df_PD_mirzapur, count_df_TOD_mirzapur))  %>% mutate(Site = 'Mirzapur')
df_mir = melt(rbind(count_df_PD_mirpur %>% 
                      rename(Megasphaera= `ASV15 (Megasphaera)`) %>%
                      select(sample, Megasphaera) %>% 
                      mutate(time = 'pre-detection'), 
                    count_df_TOD_mirpur %>% select(-symptomatic_use,-Parabacteroides,-`symptomatic?`) %>% 
                      mutate(time = 'time-of-detection'))) %>% mutate(Site = 'Mirpur')
df = rbind(df_mirz,df_mir)

```

# Save data for other scripts

```{r}

save(list = c("df","ps_mirzapur_rare"), file = paste0(paste0("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P5output_plot_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

```

# Statistics

```{r}

t.test(df$value[df$Site == 'Mirpur'], df$value[df$Site == 'Mirzapur'])
df = data.frame(Site = c('Mirzapur','Mirpur'),
  Megasphaera = c(sum(count_df_PD_mirzapur$Megasphaera>0.2),sum(count_df_PD_mirpur$Megasphaera>0.2)),
  Total = c(length(count_df_PD_mirzapur$Megasphaera),length(count_df_PD_mirpur$Megasphaera))) %>% 
  mutate(NO_mega = Total - Megasphaera) %>% select(-Total)
rownames(df) = df$Site; df$Site = NULL
chisq.test(df)

```

# Make figures

```{r}

#rmarkdown::render("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/code/R_markdown/P5_mirzapur_analysis_figures.Rmd")

```