### 2021-02-22 code/R_markdown/P2_cohort_figures.Rmd
figure mods for response to reviewers

---
title: "Microbiome analysis - cohort figures"
author: "Maureen A. Carey"
date: "09/01/2020"
output: html_document
---

# Load packages and data

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(ggplot2); library(rmarkdown); library(vegan); library(dotwhisker)
library(dplyr); library(phyloseq); library(tibble)

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000); theme_set(theme_bw())
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_plot_2020_09_11.RData")

```

## Figure 2 - Emphasizing amount of diarrhea and antibiotic usage

Figure 2A
``` {r}

df_plot = df_fig2A
colors_temp = colors; names(colors_temp)[names(colors_temp) == 'no crypto'] = "not in microbiome cohort"
colors_temp_use = colors_temp; colors_temp_use[names(colors_temp_use) != 'not in microbiome cohort'] = c("red","red","red")

p = ggplot(data = df_plot) +
  geom_histogram( aes(x = n_diarr, fill = TYPE), bins = 45)+
  scale_fill_manual(values = colors_temp_use) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.text.y = element_text(size=12),
        axis.title.x = element_text(size=20),
        axis.title.y = element_text(size=20),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.position = c(.6,.7),
        legend.title = element_blank()) + xlab('episodes of diarrhea') +
  guides(fill = F) +
  ylab('number of children')
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig2A_diarrhea_freq_mirpur.png",
       p, width = 4, height = 4, dpi = 600)
rm(p, df_plot, colors_temp, colors_temp_use)

```

Supplemental Figure 2A
```{r}

n_kids = length(unique(df_sf2A$CHILDID))
df_sf2A %>% 
  ggplot(aes(x = age, y = (..count..)/n_kids)) +
  stat_bin(binwidth = 30) +
  geom_bar() + facet_wrap(~value_use)+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm")) +
  geom_vline(xintercept = 365, linetype = "dotted") +
  ylab('percent of children') + xlab('age in days') + scale_y_continuous(labels = scales::percent)
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Sfig2A_antibiotic_usage_mirpur.png",
       width = 6.8, height = 5,units = "in",dpi = 300)

```

Figure 2B
```{r}

# stats and plot
my_comparisons <- list(#c("no infection\ndetected", "subclinical"), 
                       c("no infection\ndetected", "both"),
                       c("subclinical","both"),
                       c("subclinical", "diarrheal"))
colors_temp = colors; colors_temp['no infection\ndetected'] = "#666666"
p = ggplot(data = df_fig2B, aes(x = TYPE, y = n, color = TYPE)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.1, position = position_jitter(width = 0.3, height = 0.01, seed = SEED_INT)) +
  scale_color_manual(values = colors_temp) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust = 1),
        plot.margin=unit(c(.25,.5,.5,.25),"cm")) + 
  ggpubr::stat_compare_means(comparisons = my_comparisons, size = 2, method = 't.test') +
  ylab('episodes of diarrhea (age 0-2yrs)') + xlab(NULL) + guides(color = F)
# SOMETHING IS WEIRD -> NUMBER OF DIARRHEAL POINTS IS TOO LOW - NEVERMIND \/
# this is not using the microbiome samples' definition of subclin/diarrhea, this is using the FULL dataset
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig2B_num_diarrheal_events_mirpur.png",
       p , width =2, height = 3,units = "in",dpi = 300)

```

Figure 2C
```{r}

# stats and plot
my_comparisons <- list(#c("no infection\ndetected", "subclinical"), 
                       #c("no infection\ndetected", "both"), 
                       #c("subclinical", "both"), 
                       #c("diarrheal", "both"), 
                       c("no infection\ndetected", "diarrheal"))
                       #c("subclinical", "diarrheal"))
colors_temp = colors; colors_temp['no infection\ndetected'] = "#666666"
p = ggplot(data = df_fig2C, aes(x = TYPE, y = n, color = TYPE)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.1, position = position_jitter(width = 0.3, height = 0.01, seed = SEED_INT)) +
  scale_color_manual(values = colors_temp) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust = 1),
        plot.margin=unit(c(.25,.5,.5,.25),"cm")) + 
  ggpubr::stat_compare_means(comparisons = my_comparisons, size = 2, method = 't.test') +
  ylab('doses of antibiotics (age 0-2yrs)') + xlab(NULL) + guides(color = F)
# this is not using the microbiome samples' definition of subclin/diarrhea, this is using the FULL dataset
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig2C_num_abx_events_mirpur.png",
       p , width =2, height = 3,units = "in",dpi = 300)

```

## Figure 3

Frequent reads overall - Figure 3A + 3D
```{r}

rare_min = 10000
p1 = ggplot(df_fig3A, 
           aes(x = names, y = nreads/(rare_min*nrow(sample_data(ps_for_explore_fig3A))), fill=genus)) + geom_bar(stat="identity")
p1 = p1 + theme(axis.text.x = element_text(size=14,angle=45,hjust=1),
              axis.text.y = element_text(size=14),
              axis.title.x = element_text(size=20),
              axis.title.y = element_text(size=20),
              plot.margin=unit(c(.5,.5,.5,.5),"cm")) + 
  labs(y="Fraction of all\nreads from ASV", x="ASV assigned genus") +
  scale_x_discrete(labels = df_fig3A$genus) + 
  scale_fill_grey() + guides(fill = F) + scale_y_continuous(labels = scales::percent)
p1
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig3A_abundant_asvs_mirpur.png",
       plot=p1, width = 6.8,height = 4.2,units = "in",dpi = 300)

## to show lack of dominance of individual microbes

p_alt = ggplot(alt_df, aes(x = reorder(sample, percent_reads), y = percent_reads, fill = Genus)) +
  geom_bar(stat = "identity")
p_alt = p_alt + theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              axis.text.y = element_text(size=14),
              axis.title.x = element_text(size=18),
              axis.title.y = element_text(size=18),
              plot.margin=unit(c(.7,.5,.5,.5),"cm")) + 
  scale_fill_grey(na.value = "grey") +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
  labs(y="Percent of sample reads\nfrom most common ASV", x="samples") #+ guides(fill = F)

symptomatic.lab <- c("diarrheal", "subclinical"); names(symptomatic.lab) <- c("TRUE", "FALSE")
pre_detection.lab <- c("pre-detection", "time-of-detection"); names(pre_detection.lab) <- c("TRUE", "FALSE")
p = ggplot(alt_df, aes(x = reorder(sample, percent_reads), y = percent_reads)) +
  geom_bar(stat = "identity") + 
  facet_grid(symptomatic~pre_detection,labeller = labeller(
    symptomatic = symptomatic.lab, pre_detection = pre_detection.lab)) 
p = p + theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              axis.text.y = element_text(size=14),
              axis.title.x = element_text(size=18),
              axis.title.y = element_text(size=18),
              plot.margin=unit(c(.7,.5,.5,.5),"cm")) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + scale_fill_grey(na.value = "grey") +
  labs(y="Percent of sample reads\nfrom most common ASV", x="samples") #+ guides(fill = F)
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig3_dominant_asvs.png", 
       plot=p, width=6.8, height = 4.2,units = "in",dpi = 300)

p = ggplot(df_fig3D, aes(x = names, y = nsamples/(nrow(sample_data(ps_for_explore_fig3A))), fill=genus)) + geom_bar(stat="identity")
p = p + theme(axis.text.x = element_text(size=14,angle=45,hjust=1),
              axis.text.y = element_text(size=14),
              axis.title.x = element_text(size=18),
              axis.title.y = element_text(size=18),
              plot.margin=unit(c(.7,.5,.5,.5),"cm")) + 
  labs(y="Percent of stool samples\ncontaining ASV", x="genus of ASV") + guides(fill = F) +
  scale_x_discrete(labels = df_fig3D$genus) + scale_fill_grey(na.value = "grey") + xlab(NULL) +
  geom_hline(yintercept = .5, linetype="dashed", 
                color = "gray", size=1) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1))
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig3D_samplecount_asvs_mirpur.png",
       plot=p, width=6.8, height = 4.2,units = "in",dpi = 300)

p = ggplot(alt_countsumsdf, aes(x = names, y = nsamples/(nrow(sample_data(ps_rare_bygenus))), fill=genus)) + geom_bar(stat="identity")
p = p + theme(axis.text.x = element_text(angle=45,hjust=1),
              axis.text.y = element_text(size=12),
              axis.title.x = element_text(size=20),
              axis.title.y = element_text(size=20),
              plot.margin=unit(c(.5,.5,.5,.5),"cm")) + 
  labs(y="Fraction of samples containing ASV", x="ASV assigned genus") +
  scale_x_discrete(labels = alt_countsumsdf$genus) + guides(fill = F) + #readsumsdf?
  scale_y_continuous(labels = scales::percent, limits = c(0,1))

```

Richness and Evenness - Figure 3B and 3C

```{r}

theme_set(theme_bw())
p_observed = plot_richness(ps_for_alpha,x='Clinical.type', measures = "Observed", color = 'Clinical.type') + 
  geom_boxplot(alpha=0.5, outlier.shape = NA) + 
  geom_point(size = 1,alpha = 0.25,position = position_jitter(w = 0.1, h = 0)) + 
  labs(y="Richness (observed ASVs)",x=NULL) + 
  theme(strip.text.x=element_blank(),
        axis.title.x = element_text(size=20), 
        axis.title.y = element_text(size=20),
        axis.text.x = element_text(size=12,angle=-45,hjust=0,vjust = 1),
        plot.margin=unit(c(.5,.5,.5,.5),"cm")) +
  scale_x_discrete(labels = c('prior to\nsubclinical', 'prior to\ndiarrheal', 'subclinical','diarrheal')) +
  scale_color_manual(values = c("#666666","#666666","#FF7F00","#1F78B4")) +
  guides(color = F)
p_observed$layers <- p_observed$layers[-1]
p_observed
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig3B_Observed_div_mirpur.png",
       plot=p_observed,width=4.5, height = 4.2,units = "in",dpi = 300)

p_simpson = plot_richness(ps_for_alpha,measures = "InvSimpson", x='Clinical.type', color = 'Clinical.type') + 
  geom_boxplot(alpha=0.5, outlier.shape = NA) + 
  geom_point(size = 1,alpha = 0.25,position = position_jitter(w = 0.1, h = 0)) + 
  labs(y="Evenness (Inv. Simpson)",x=NULL) + 
  theme(strip.text=element_blank(),
        axis.title.x = element_text(size=20), 
        axis.title.y = element_text(size=20),
        axis.text.x = element_text(size=12,angle=-45,hjust=0,vjust = 1),
        plot.margin=unit(c(.5,.5,.5,.5),"cm")) +
  scale_x_discrete(labels = c('prior to\nsubclinical', 'prior to\ndiarrheal', 'subclinical','diarrheal')) +
  scale_color_manual(values = c("#666666","#666666","#FF7F00","#1F78B4")) +
  guides(color = F)
p_simpson$layers <- p_simpson$layers[-1]
p_simpson
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig3C_invsimpson_mirpur.png",
       plot=p_simpson, width=4.5, height = 4.2,units = "in",dpi = 300)

```

## Figure 4A

```{r}

p = plot_ordination(
  physeq = ps_rare_for_beta,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "Number.of.crypto.positive.events"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "infection", shape = "Number of infections")
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4A_pcoa_mirpur_euclidean.png", 
       plot=p, width = 5.4,height = 3,units = "in",dpi = 600)

```

```{r}

ps_rare_for_beta_mode_sample_data = ps_rare_for_beta
qu = quantile(sample_data(ps_rare_for_beta_mode_sample_data)$Age.in.days, c(0.25, 0.5, 0.75))
temp_sample_data = as(sample_data(ps_rare_for_beta_mode_sample_data),"data.frame") %>% 
  rownames_to_column('sample') %>%
  mutate(age_quartile = as.factor(ifelse(Age.in.days < qu[[1]][1], 1,
                               ifelse(Age.in.days > qu[[3]][1], 4,
                                      ifelse(Age.in.days > qu[[2]][1], 3,2)))),
         hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)),
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>%
    column_to_rownames('sample')
sample_data(ps_rare_for_beta_mode_sample_data) = temp_sample_data
p = plot_ordination(
  physeq = ps_rare_for_beta_mode_sample_data,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "age_quartile"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  #scale_shape_manual(values = c(5,16,4),
   #                  labels = c("1" = "youngest quartile","2" = "middle","4" = "oldest quartile") )+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "symptomatic?", shape = "Age (quartitled)")
p

p = plot_ordination(
  physeq = ps_rare_for_beta_mode_sample_data,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "meleagridis"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  scale_shape_manual(values = c(5,16),
                     labels = c("1" = "positive","0" = "negative") )+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "symptomatic?", shape = "C. meleagridis")
p

p = plot_ordination(
  physeq = ps_rare_for_beta_mode_sample_data,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "hominis"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  scale_shape_manual(values = c(5,16),
                     labels = c("1" = "positive","0" = "negative") )+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "symptomatic?", shape = "C. hominis")
p

```

Do all kids have at least one abx between predetction and time of detection???

```{r}

ggplot(df_abx_between, aes(x = abx_between_samples, y = value, fill = symptomatic)) + geom_bar(stat = "identity",position = "dodge")+
  annotate(geom = "text", x = 1.8, y = 70, label = "Pearson's Chi-\nsquared test\np-value = NS") +
  scale_fill_manual(values= c(colors['diarrheal'],colors['subclinical'])) +
  ylab('number of children') + xlab('antibiotics between pre-detection\nand time-of-detection samples?') +
  guides(fill= guide_legend(title='symptomatic?'))

```

Do sympotmatic kids have more abx before time_of_detection sample?

```{r}

my_comparisons = list(c("subclinical","diarrheal"))
ggplot(df_abx_total,aes(x= symptomatic, y = n_abx_doses, color = symptomatic)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitter(width = 0.3, height = 0, seed = SEED_INT)) +
  scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) +
  ylab('number of antibiotic doses') + xlab('symptomatic?') + guides(color = F) +
  ggpubr::stat_compare_means(comparisons = my_comparisons, size = 3, method = 't.test')

```

Do sympotmatic kids have more abx before time_of_detection sample MONTH PRIOR ONLY?

```{r}

my_comparisons = list(c("subclinical","diarrheal"))
ggplot(df_abx_month,aes(x= symptomatic, y = n_abx_doses, color = symptomatic)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitter(width = 0.3, height = 0, seed = SEED_INT)) +
  scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) +
  ylab('number of antibiotic doses') + xlab('infection') + guides(color = F) +
  ggpubr::stat_compare_means(comparisons = my_comparisons, size = 3, method = 't.test')
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig2B_abx_MONTH_mirpur.png", 
              width = 3, height = 4, dpi = 600)

```

What happens if we break this down to look at a single drug at a time?

```{r}

ggplot(df_individual_abx, aes(x = value,y = variable, color = symptomatic)) + 
  geom_boxplot(position = "dodge", outlier.shape = NA) + 
  geom_point(alpha = 0.1, 
             position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0, seed = SEED_INT, dodge.width = 0.75)) +
  scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) +
  ylab('Antibiotics') + xlab('number of doses') +
  guides(color= guide_legend(title='symptomatic?')) +
  theme(legend.position = c(.8,.15))

```

Also check number of distinct antibiotics

```{r}

p = ggplot(df_distinct_abx, aes(x = n_distinct_abx,y = symptomatic, color = symptomatic)) + 
  geom_boxplot(position = "dodge", outlier.colour = NA) + geom_point(alpha = 0.2, position = position_jitter(width = 0.3, height = 0.3, seed = SEED_INT)) +
  scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) + guides(color = F) +
  xlab('number of distinct antibiotics used') + ylab('symptomatic?')
p

dwplot(abx_model, vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) %>%
  relabel_predictors(c(n_distinct_abx = "number of distinct antibiotics",
                       Nalidixic_acid = "Nalidixic acid")) +
     theme_bw() + xlab("Coefficient Estimate") + ylab("") +
  theme(plot.title = element_text(face="bold"),
        plot.margin = unit(c(1.5,1,1,1), "cm")) + 
  scale_color_manual(values = c("black")) + 
  guides(color = F)

```